---
title: "Project"
author: "Jacob Yousif"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setup

Libraries used:
```{r}
library(tidyverse)
library(janitor)
library(naniar)
library(paletteer)
```

Load data
```{r}
exoplanets <- read_csv("data/cleaned_5250.csv")
planets_temps <- read_csv("data/oec.csv")
```

Explore data
```{r}
glimpse(exoplanets)
miss_var_summary(exoplanets)
```

Add column for mass of planets in kg
```{r}
exoplanets_mass <- exoplanets %>% 
  mutate(mass_kg = case_when(mass_wrt == "Jupiter" ~ mass_multiplier * 1.898e27, #mass of Jupiter
                             mass_wrt == "Earth" ~ mass_multiplier * 5.972e24)) #mass of Earth
glimpse(exoplanets_mass)
```
```{r}
space_palette <- paletteer_d("palettetown::swampert")
space_palette2 <- paletteer_d("palettetown::swampert", 10, direction = -1)
space_palette_cont <- paletteer_d("palettetown::swampert", type = "continuous")
space_palette
space_palette2
space_palette_cont
```

Discovery year vs size
```{r}
ggplot(exoplanets_mass, aes(x=discovery_year, 
                            y=mass_kg, 
                            group = discovery_year, 
                            fill = cut(discovery_year, breaks = 
                                         c(1990, 1995, 2000, 2005, 2010, 2015, 2020, 2023)
                                       ))) +
  geom_boxplot(na.rm=T) +
  scale_y_log10() +
  scale_fill_manual(values = space_palette) +
  labs(title="Mass of Planets Discovered Each Year", x="Discovery Year", y="Mass (kg)", fill="Year Discovered")
```
Discovery year vs distance
```{r}
ggplot(exoplanets_mass, aes(x=discovery_year, 
                            y=distance, 
                            group = discovery_year, 
                            fill = cut(discovery_year, breaks = 
                                         c(1990, 1995, 2000, 2005, 2010, 2015, 2020, 2023)
                                       ))) +
  geom_boxplot(na.rm=T) +
  scale_y_log10() +
  scale_fill_manual(values = space_palette2) +
  labs(title="Distance of Planets Discovered Each Year", x="Discovery Year", y="Distance from Earth (ly)", fill="Year Discovered")
```

Size vs eccentricity
```{r}
exoplanets_mass %>% 
  filter(eccentricity >= 0.0) %>% #why are some ecc negative?
  ggplot(aes(x=mass_kg, y=eccentricity)) +
  geom_point() +
  scale_x_log10()
```

```{r}
#unreactive version:
planets_temps %>% 
  select("PlanetIdentifier", "PlanetaryMassJpt", "Eccentricity", "PeriodDays", "SurfaceTempK") %>% 
  filter(PlanetIdentifier == "Earth" | PlanetIdentifier == "Kepler-9 b")# %>% 
  #ggplot(aes(x = PlanetIdentifier, y = PlanetaryMassJpt, group = PlanetIdentifier)) +
  #    geom_col(position = "dodge")
```
```{r}
return(TRUE)
```

```{r}
planets_temps %>% 
  select(PlanetIdentifier, PlanetaryMassJpt, Eccentricity, PeriodDays, SurfaceTempK) %>% 
  filter(PlanetIdentifier == "Earth" | PlanetIdentifier == input$planet)
  
```


```{r}
library(shiny)
planets_temps <- read_csv("data/oec.csv")

ui <- fluidPage(
  title = "Compare A Planet with Earth",
  selectInput("planet", "Select Exoplanet...", choices = NULL, 
              selected = "Earth", width = "400px"),
  
  fluidRow(
    plotOutput("plot1", width = "300px", height = "500px"),
    plotOutput("plot2", width = "300px", height = "500px"),
    plotOutput("plot3", width = "300px", height = "500px")
  #plotOutput("plot4", width = "500px", height = "500px")
  )
)

server <- function(input, output, session) {
  
  updateSelectInput(session, "planet", choices = planets_temps$PlanetIdentifier) #had to use this because selectInput had a large number of choices
  
  #debug:
  #observe({print(input$planet)}) #indicates whether selectInput is working
  
  get_planet_data <- reactive({ #this should work
    planets_temps %>% 
      select(PlanetIdentifier, PlanetaryMassJpt, Eccentricity, PeriodDays, SurfaceTempK) %>% 
      filter(PlanetIdentifier == "Earth" | PlanetIdentifier == input$planet)
                      }) 

  output$plot1 <- renderPlot({ #mass
    ggplot(data = get_planet_data(), aes(x = PlanetIdentifier, y = PlanetaryMassJpt, group = PlanetIdentifier)) +
      geom_col(position = "dodge")
  })
  
  output$plot2 <- renderPlot({ #Eccentricity
    ggplot(data = get_planet_data(), aes(x = PlanetIdentifier, y = Eccentricity, group = PlanetIdentifier)) +
      geom_col(position = "dodge")
  })
  
  output$plot3 <- renderPlot({ #Orbital period
    ggplot(data = get_planet_data(), aes(x = PlanetIdentifier, y = PeriodDays, group = PlanetIdentifier)) +
      geom_col(position = "dodge")
  })
  
  #output$plot4 <- renderPlot({ #Surface temp: no worky cuz earth has na
  #  ggplot(data = get_planet_data(), aes(x = PlanetIdentifier, y = SurfaceTempK, group = PlanetIdentifier)) +
  #    geom_col(position = "dodge")
  #})
  
  
  #Stylize plots
}

shinyApp(ui, server)
```

Idea: exoplanet habitability calculator
- Select exoplanet by name?
- Or, maybe sort list of viable exoplanets by a metric (like distance, similarity to Earth in size/day cycle/temp/year length)

Categories: 
- Size (relative to Earth)
- Distance from Earth
- Time it would take to reach with current tech
- Surface temp range
- Day length/year length
- Eccentricity (could you make a vis of eccentricity w/ graphs?)
- How long until planet is consumed? (based on star size/type)